---
title: "Homework 1"
author: "Shun-Lung Chang, Dilip Hiremath"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", fig.align = 'center', fig.height = 4, fig.width = 4)
```

```{r, include=FALSE}
library(knitr)
```

## 1. Before you address your real task you just want to get a bit acquainted with the data. Read the data into R and use the function summary to get a first overview on the data.

```{r}
load("data/Forbes2016.Rdata")
summary(Forbes2016)
```

### (a) How many companies in the data set are located in China?

There are `r nrow(Forbes2016[Forbes2016$Country == "China", ])` companies located in China, as one can see from the summary table. 

### (b) How many companies in the data set have missing values (i.e. are listed as NA)? For which variables are data missing?

If we look at *Assets* and *Market.Value* variables, there is 1 missing value in both variables. The row containing missing values is shown below.

```{r}
kable(Forbes2016[is.na(Forbes2016$Assets), ])
```

```{r}
kable(Forbes2016[is.na(Forbes2016$Market.Value), ])
```

In addition, there exist four rows containing missing values in *Sales* variable. These rows are shown below.

```{r}
kable(Forbes2016[is.na(Forbes2016$Sales), ])
```

### (c) Which four countries have the top most companies in the Forbes’ list?

The first top countries are United States, Japan, China, and United Kingdom.

### (d) Which share of companies in the data set have a market value smaller than or equal to 6.30 billion USD?

The share is `r format(sum(Forbes2016$Market.Value <= 6.3, na.rm = TRUE) / nrow(Forbes2016) * 100, digits = 4)`%, and the value can be counted in the following code.

```{r}
sum(Forbes2016$Market.Value <= 6.3, na.rm = TRUE) / nrow(Forbes2016)
```

### (e) Which share of companies in the data set have assets that are larger than or equal to 10.40 billion USD?

The share is `r format(sum(Forbes2016$Assets >= 10.4, na.rm = TRUE) / nrow(Forbes2016) * 100, digits = 4)`%, and the value can be counted in the following code.

```{r}
sum(Forbes2016$Assets >= 10.4, na.rm = TRUE) / nrow(Forbes2016)
```

## 2. Let us have a closer look at the variable types involved here. Looking at the content of each variable:

### (a) Which of the variables in the data set are discrete? 

The discrete variables are *Rank*, *Company*, and *Country*.

### (b) Which of the variables are continuous?

The discrete variables are *Sales*, *Profits*, *Assets*, and *Market.Value*.

### (c) Which discrete ones can be ordered?

The *Rank* variable can be orderd.

## 3. Which types are used in R to store the variables in the data frame? 

The following code returns the class of each variable.

```{r}
sapply(Forbes2016, class)
```

## 4. Look at the summary statistics to see how the variable class affects the output. Which variable classes produce the same summary output? Which produce a different one?

The variables belonging to class **integer** or **numeric** produce the measures of center and dispersion. On the other hand, variables in class **factor** produce the counts of each class.

## 5. Draw a histogram for the variable Profit using the default number of bins and color the bins in 'springgreen'. Add a meaningful title to the plot.

```{r}
hist(Forbes2016$Profits, col = "springgreen",
     main = "Counts of Profits", xlab = "Profits")
```

### (a) Is the distribution of market values in the data set symmetric?

As can been from the plots, both the distributions of *Profits* and *Market.Value* are not symmetric.

```{r}
hist(Forbes2016$Market.Value, col = "springgreen", 
     main = "Counts of Market Values", xlab = "Market Values")
```

### (b) Does the selected number of bins yield a good graphical representation of the data? Try out 7, 50 and 100 bins! Is any of these three options more suited?

In fact, there is no optimal number for bins. But in this case, we would say that 50 is better since the bins are not too small or too large.

```{r}
hist(Forbes2016$Profits, col = "springgreen", breaks = 7, 
     main = "Counts of Profits (bins = 7)", xlab = "Profits")
```

```{r}
hist(Forbes2016$Profits, col = "springgreen", breaks = 50, 
     main = "Counts of Profits (bins = 50)", xlab = "Profits")
```

```{r}
hist(Forbes2016$Profits, col = "springgreen", breaks = 100, 
     main = "Counts of Profits (bins = 100)", xlab = "Profits")
```

## 6. Draw a boxplot for the variable Profit.

```{r}
boxplot(Forbes2016$Profits, horizontal = TRUE, main = "Boxplot of Profits")
```

### (a) Compare the boxplot with the histograms above.

### (b) Which graphical representation of the data do you find more helpful?

### (c) How many companies do have profits above 40 billion USD? Can you find out which companies these are?

```{r}
Forbes2016[Forbes2016$Profits >= 40, ]$Company
```

## 7. You have a closer look at profits of the companies.

### (a) First, you want to know which profit is needed such that a company belongs to the top two percent. 

```{r}
quantile(Forbes2016$Profits, probs = 0.98)
```

### (b) Which percentage of companies has negative profits?

```{r}
sum(Forbes2016$Profits < 0) / nrow(Forbes2016)
```

### (c) How large is the interquartile range in sales for the companies on Forbes’ list?

```{r}
IQR(Forbes2016$Sales, na.rm = TRUE)
```

## 8. Now you start turning towards your real task. Create a subset of the data set comprising all companies that have their legal residence in one of the following countries: Brazil, India, South Korea, France, Germany or Mexico. Draw a boxplot for the variable Profits splitted by country for these companies.

```{r, fig.height=5}
sub_forbes2016 <- subset(Forbes2016,
                         Country %in% c("Brazil", "India", "South Korea", "Franc", 
                                        "Germany", "Mexico"))

boxplot(Profits ~ as.character(Country), data = sub_forbes2016, horizontal = TRUE, 
        main = "Boxplot of Profits in different Countries",
        las = 1)
```

### (a) How much difference in medians between countries is shown in the plot?

### (b) Which country shows the largest median profit?

The plot indicates that the largest median profit lies in Brazil. 

### (c) Which country/countries show the most symmetric distributions for profit?

South Korea shows the most symmetric distributions for profit in the plot.

### (d) Create a frequency table for the variable Country sorted in decreasing order. Which three countries – among those selected – have the top most companies in Forbes’ list?

```{r}
sort(table(as.character(sub_forbes2016$Country)), decreasing = TRUE)
```

## 9. Now you start with your real task for the subset data created in Question 8.

### (a) What is the median profit of the selected companies?

```{r}
median(sub_forbes2016$Profits, na.rm = TRUE)
```

### (b) What is the average profit of the selected companies?

```{r}
mean(sub_forbes2016$Profits, na.rm = TRUE)
```

### (c) What is the average distance for a company’s profit to the average profit (i.e. what is the standard deviation of profits)?

```{r}
sd(sub_forbes2016$Profits, na.rm = TRUE)
```

### (d) Does the top third of selected companies have profits that are at least as high as the median profit for all companies on the Forbes list?

```{r}
aggregate(sub_forbes2016$Profits, list(sub_forbes2016$Country), median)
```

## 10. In order to fulfil your task you finally draw a random sample of size 80 from the list of your selected companies. Create the resulting data frame!

```{r}
set.seed(42)
sampled_forbes2016 <- sub_forbes2016[sample(1:nrow(sub_forbes2016), 80), ]
```

### (a) Compute the mean profit for your sample? 

```{r}
mean(sampled_forbes2016$Profits, na.rm = TRUE)
```

### (b) Compute the median profit for your sample?

```{r}
median(sampled_forbes2016$Profits, na.rm = TRUE)
```

### (c) Compute the minimum profit for your sample? 

```{r}
min(sampled_forbes2016$Profits, na.rm = TRUE)
```

### (d) Compute the maximum profit for your sample?

```{r}
max(sampled_forbes2016$Profits, na.rm = TRUE)
```

### (e) Compute the 90%-quantile for your sample?

```{r}
quantile(sampled_forbes2016$Profits, probs = 0.9)
```

## 11. Are median and mean profits of your sample below or above the median and mean market value for all companies in the selected countries?


## 12. Compute the standard deviation of profits for your sample. Is it smaller or larger than the standard deviation in profits for all selected companies?

```{r}
sd(sampled_forbes2016$Profits, na.rm = TRUE)
```

## 13. Draw a boxplot, a histogram, and a density plot of the profit for your sample. Can you draw any substantial information from your plot? (e.g about skewness of the distribution, outliers, quartiles)

```{r}
boxplot(sampled_forbes2016$Profits, horizontal = TRUE, 
        main = "Boxplot of Profits for Sampled Data")
```

```{r}
hist(sampled_forbes2016$Profits, 
     main = "Counts of Profits for Sampled Data", xlab = "Profits")
```

```{r}
d <- density(sampled_forbes2016$Profits)
plot(d, main = "Density of Profits for Sampled Data", xlab = "Profits")
```
