---
title: "Homework 5"
author: "Shun-Lung Chang, Dilip Hiremath"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, comment = "", fig.align = 'center', fig.height = 3, fig.width = 6, message = FALSE, warning = FALSE)
```

```{r}
library(magrittr)
library(dplyr)
library(ggplot2)
library(car)
```

## 1. First of all, load the data frame Wage from the library ISLR. You start out with a close look at wage differences between the two health levels.

```{r}
data(Wage, package = "ISLR")
```

### (a) (1.5 points) Compute mean and standard deviation of wage for each health level separately. Summarize the result in an English sentence.

```{r}
wage_stats <- Wage %>% 
    group_by(health) %>% 
    summarise(mean_wage = mean(wage),
              sd_wage = sd(wage),
              counts = n())
wage_stats
```
There are 858 observation for good health or below workers and their average wage is 101.6613 with the standard deviation of 35.18500. There are 2142 observation for very good health or above workers and their average wage is 115.7262 with the standard deviation of 43.43896.

### (b) (1 point) Compute the standard errors for the mean wages in the two groups.

```{r}
wage_stats$se_wage <- wage_stats$sd_wage / sqrt(wage_stats$counts)
wage_stats
```
The standard errors for the mean wage for good health or below workers is 1.2011960 and for very good health or above workers is 0.9385766.

## 2. (2.5 points) Create a plot showing the mean wages for the two groups and corresponding error bars, i.e. add lines of length one standard error of the mean to both sides of the mean.

```{r}
ggplot(wage_stats, aes(x = health, y = mean_wage, group = 1)) +
    geom_errorbar(aes(x = health, ymin = mean_wage - se_wage, ymax = mean_wage + se_wage), 
                  width = 0.1) +
    geom_point() +
    labs(x = 'Health' , y = 'Wage')
```
The plot is as show above.

## 3. (2.5 points) Using an appropriate statistical procedure, test whether average wage is the same for workers with health level "1. at most Good" and workers with health level "2. at least Very Good". Formulate the null and alternative hypothesis and report the results in an English sentence refering to the relevant numbers.

$H_0 : Average\ wage_{level_1} = Average\ wage_{level_2}$

$H_1 : Average\ wage_{level_1} \neq Average\ wage_{level_2}$ 

```{r}
# Assume that population variances of the two classes are not equal
t.test(Wage$wage ~ Wage$health, var.equal = FALSE)
```
The null hypothesis is that the average wage for level 1 at most Good is equal to average income of workers with health level 2 at least Very Good.
The alternative hypothesis states that the average wage for level 1 at most Good is not the same as average income of workers with health level 2 at least Very Good.
we see that based on the test results we can reject the null hypothesis in favour of the alternative hypothesis. 
The difference in mean wages between level 1 at most Good (101.6613) and average income of workers with health level 2 (115.7262) is significant  in a  two-sided, Welch-test(t = -9.2265,  df = 1934.3, p-value < 2.2e-16).

## 4. Plot a box plot of the workers raw wage (variable wage using the education level (variable education) as grouping.

```{r}
ggplot(Wage) +
    geom_boxplot(aes(x = education, y = wage))
```

### (a) (half a point) Are half of the wages for workers who have less than a high school degree below the first quartile of the wage for workers with some college degree?
No

### (b) (half a point) Do half of the workers with a HS degree have higher wages than three quarters of the high school dropouts?
No

### (c) (half a point) The minimum wage of workers with advanced degree is larger than the median wage of high school dropouts?
No ??????????????????????Kindly check this ???????????????????????????????????????????????????????????????????

### (d) (half a point) The interquartile range differs substantially between all groups.
No  ??????????????????????Kindly check this ???????????????????????????????????????????????????????????????????

### (e) (half a point) Spread as measured by the length of the whiskers differs substantially between all groups.
No ??????????????????????Kindly check this ???????????????????????????????????????????????????????????????????

## 5. You want to assess the wage difference between educational groups. Before you run the appropriate statistical test, you check some of the assumptions for ANOVA. In particular, you assess homoscedasticity.

### (a) (1.5 points) Looking at the boxplot in Question 4. Does homoscedasticity hold for the five groups? Give reasons for your answer!
No

### (b) (1 point) Select a suitable variance test to check on this. Does the test confirm homoscedasticity?

```{r}
leveneTest(Wage$wage, Wage$education)
```
No, this test does not confirm homoscedasticity. Which is in line with out analysis from the box plot. We observe that the P value is very less and hene we can reject the null hypothesis in favour of the alternative.

## 6. Now, you assess the wage difference between educational groups using a statistical test.

### (a) (1 point) Using an appropriate statistical test check whether wages are equal across education groups. Report the result in a complete English sentence including the relevant numbers!

```{r}
wage_education <- aov(wage ~ education, data = Wage)
anova(wage_education)
```
There is a highly significant difference in wages across educational groups as given by the ANOVA test with a test-statistic of F = 229.81 with 4 numerator
and 2995 denominator degrees of freedom yielding a p-value of p < 2.2e-16.

### (b) (1 point) From the ANOVA table derive the total sum of squares for wages and compare this result with the variance of wage when multiplied by 2999.

```{r}
sum(anova(wage_education)[, 2])
```

```{r}
var(Wage$wage) * 2999
```
The are both equal as show in the above calculations.

### (c) (half a point) Which proportion of total variation in wages is due to the group differences in education?

```{r}
anova(wage_education)[1, 2] / sum(anova(wage_education)[, 2])
```
0.2348419 is the proportion of total variation in wages due to the group differences in education.

## 7. Having found an overall difference, you now want to use a post-hoc test with Holm correction, to asses which marital status groups do actually differ significantly in wages?

```{r}
pairwise.t.test(Wage$wage, Wage$education, p.adjust.method = "holm")
pairwise.t.test(Wage$wage, Wage$maritl, p.adjust.method = "holm")
```

### (a) (2 points) According to the post hoc test which groups differ significantly?
for education groups 
Some College - < HS Grad
College Grad - < HS Grad
Advanced Degree - < HS Grad
Some College - HS Grad
College Grad -  HS Grad
Advanced Degree - HS Grad
College Grad -  Some College
Advanced Degree - Some College
College Grad - Advanced Degree

for Martial groups
never married - married
never married - divorced
married - divorced
married - separated


###(b) (half a point) According to the post hoc test which groups do not differ significantly?

for education groups 
HS Grad - < HS Grad

for Martial groups
never married - widowed
never married - separated
married - widowed
widowed - divorced
widowed - separated
divorced - separated

## 8. You now investigate the relationship between wage and the two predictors education and health status.

### (a) (1 point) First, calculate a main effects model only. Give a verbal summary of the model result!

```{r}
wage_edu_heal <- aov(wage ~ education + health, data = Wage)
anova(wage_edu_heal)
```
There is a highly significant difference in wages across education groups as given by the ANOVA test with a test-statistic of F =  231.248 with 4 numerator
and 2995 denominator degrees of freedom yielding a p-value of < 2.2e-16.
There is a difference but it is not too significant in wages across health groups as given by the ANOVA test with a test-statistic of F =  19.791 with 1 numerator
and 2995 denominator degrees of freedom yielding a p-value of 8.956e-06.

### (b) (1 point) Second, calculate a model with interaction. Give a verbal summary of the model result!

```{r}
wage_edu_heal_inter <- aov(wage ~ education + health + education:health, data = Wage)
anova(wage_edu_heal_inter)
```
The difference in wages caused by the combined effect of health and educational groups is not significant as given by the ANOVA test with a test-statistic of F =  1.2316 with 4 numerator and 2995 denominator degrees of freedom yielding a p-value of 0.2952.


### (c) (half a point) Using the TukeyHSD post-hoc tests, which education levels do actually differ significantly in wages?

```{r}
TukeyHSD(wage_edu_heal)
```
